Considerations for the Next Iteration of the DOE Fast Forward Storage and IO Project

### 2014-02-26 paper.tex
Major editing pass. a quarter page too long now. One section still needs to be reconciled.

\begin{abstract}
The Fast Forward effort to define a next generation filesystem has made
admirable contributions in architecture and design. Formalizing the general
idea of data staging as burst buffers for the file system will help manage
the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a tranasctional mechanism to
manage faults and data visilibilty helping to enable effective analytics
without having to work around the IO stack semantics.

While these and other contributions are valuable, improvements can be
incorporated. For example, the Doubly Distributed Transactions (\DDTns)
protocol offers an alternative approach for incorporating transactional
semantics into the data path. The differing semantics between the two
approaches and the functionality provided offer alternatives. Further, the
lessons learned in implementing \DDT has generated insights into what the
FastForward file system could and perhaps should do, given different features.

This paper examines some of the choices made by the FastForward team and
compares them with other options and offers suggestions based on experiences.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, the US Department of Energy commissioned an effort
to develop a design and prototype for an IO stack suitable for the extreme
scale environment. This is a joint effort led by the Intel Lustre team, Los
Alamos National Laboratory, EMC, and the HDF Group. This team has developed a
specification for a future IO stack to address the identified challenges.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library. Below that is an IO
forwarding layer that redirects IO calls from the compute nodes to the IO
dispatching layer. This IO forwarding layer is analogous to the function of the
IO nodes in a BlueGene machine. The IO dispatcher (IOD) has considerable
functionality.

The IOD serves as the sole interface between an application and the persistent
storage array. The core idea for IOD is to provide a way to manage the IO load
that is separate from the compute nodes and the storage array. Communication
intensive acitivies, such as two-phase, data sieving IO's data rearrangement
phase, can be moved to the IOD layer offloading the communication load from the
compute nodes. IOD has three main purposes. First, if the optional burst
buffer is availble, it works as a fast cache absorbing write write operations
for the slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations. Second, it offers the transaction mechanism for controlling data
set visibility prior to it being correct and complete. Third, data processing
operations can be placed in the IOD. This operations are inteded to offer data
rearrangement, filtering, and similar operations prior to it reaching the
central storage array.

This offloads the requirement for collective two-phase data sieving
on the compute nodes to reorganize data. This technique has been proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.

@@

@@

A second feature is the incorporation of a transaction like mechanism into the
IOD layer and a related epochs feature into the DAOS layer. This feature offers
a mechanism to avoid making incomplete data visible to other users and offer a
level of protection if the data can be stored on the optional SSDs.

These features aim to address both the bursty nature of HPC IO and to make the
data storage stack resilient to faults. While these features can address these
concerns, a deeper evaluation of the design suggests exploring alternatives
may improve the performance and/or functionality desired.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incoporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and offers a comparison to the \DDT system.
Section~\ref{sec:optimizations} discusses some of the optimizations offered
by the current design and how and when these optimizations will work best. It
also explores some of the dark corners that may yield performance issues with
the current design. Finally, Section~\ref{sec:summary} disucsses the system
overall with recommendations on what design elements should be reconsidered
based on broader issues with current HPC data centers.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}
and later incorporated as part of the IO
stack~\cite{bent:2012:challenges,bent:2012:burst-buffer}. The proposed optional
use of SSDs is problematic. First, with these burst buffers being optional, the
semantics of how IOD works must change to use DAOS to store the data directly
rather than storing them locally until explicitly persisted by the user. How
this work work, the impact on metadata, and the ability to support
functionality such as data reordering, such as chainging the fast dimension of
an array, is undefined. How doing function shipping into an IOD without a burst
buffer would work is also not explored. Further thought about how to have an
IOD layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not an output is complete
or not. At the DAOS layer, they are called epochs and represent persisted
transactions from the IOD layer. Each of these offers different functionality,
but are connected as is explained below.

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container serves to host a transaction and contains a collection of
objects. Conceptually, containers corresponds to a file in a traditional file
system. The objects in each container represent different data within a file.
The three initially defined object types are key-value store, array shard,
and blob. The easiest way to understand these types is to evaluate these from
the perspective of an HDF-5 file. The key-value store represents a collection
of attributes. The array shard represents part of a potentially
multi-dimensional array. It is refered to as a shard because it is likely a
small piece of a globally defined array. The blob represents a byte stream.
The fundamental difference between an array shard and a blob is that the array
shard has metadata identifying its portion within the global, logical space
while the blob is simply a 1-dimensional array of bytes that is not shared
across IOD nodes. Should an operation be deployed into IOD to manipulate an
array within a container, it would operate on the array shards rather than on
blobs. Given this context, the transactions come in two forms.

@@

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. Each persisted copy increments the epoch. NEED TO CORRECT
THIS INFORMATION AS IT IS WRONG.

\subsection{Metadata Management}
While a central design goal of IOD and DAOS is to eliminate the metadata server
as a core component, it was not fully accomplished.  The purpose behind the goal
is to eliminate the serialized bottleneck caused by having a centralized
metadata service. Much of the design of IOD with objects and continers does
work to eliminate these bottlenecks. However, IOD does not go fully to a no
metadata service model. Instead, the IOD layer describes a key-value metadata
service it provides. This service maintains four different items:

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. First, there are no list of containers. This
is closer to the no metadata ideal, but given the rest of the functionality,
it exports to the user level a requirement to track what containers are in the
file system. Second, given the flattening that occurs between IOD and DAOS,
the mapping from IOD objects to DAOS objects is not likely to be 1-to-1. This
is likely a list instead. Further, given the ability to re-stripe in the IOD
layer, the list is not a list of objects completely included in the IOD object,
but instead objects from which part of the data came from some listed DAOS
object.  Third, sharding IOD objects makes sense based on the compute process
connection a shard has with the complete object. Striping of an IOD object
makes sense in connection with the DAOS object(s) underneath, but striping
within the IOD layer is not discussed. Further, using striping from clients
into the IOD layer introduces a level of coordination that IOD seeks to avoid.
Fourth, given the previous items, there is ambiguity with the maximum valid
offset of the object. This could be for the shard or for the collection of
shard objects that represent a distributed variable.

The \DDT project created a simplistic model for a metadata service that has
different features that is described below and has been published about
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the ojbjects in the metadata service. While it does not address having a
file equivalence like the container concept, it does introduce a way to figure
out what is available. Stored along with this object description is a list of
the global dimensions for array objects. One observtion made as part of
creating this service is that this is insufficient for many engineering codes
that use non-regular meshes. Second, for each object, a list of the equivalent
to shards are kept with a link back to the master object and the offset and
size for each dimension for that shard. This approach is not scalable because
of the explosion in the process count. Using a striping approach like IOD does
is superior. Third, there is no mapping to another layer with different object
layout and counts.

Based on the lessons from the \DDT metadata service construction, having a
completely separate metadata service is workable. Rather than making it a
bottleneck in the IO path, it is another service that users must interact with
if they need those services. Otherwise, it is 100% out of the way. Users can
manage everything by maintaining the metadata including the list of objects
themselves. However, there are drawbacks to this 100% client-side approach.

For the IOD proposal, the service needs some adjustments to be generally
useful.  First, there needs to be a way to query the list of available
containers. Second, the maximum transaction ID for a container is mentioned as
being maintained by this metadata service, but when and how this is done is
never mentioned. Second, it should be extracted from IOD entirely and kept as
a separate service along side. This will eliminate any contention caused by
interaction with the IOD layer. This also offers a way to link with the DAOS
layer in a machine-independent way. The need for this is described below.
Third, the definition of an object needs to be made more clear. In some cases,
it seems that an object is a globally distributed object is represented by a
collection of shards. In other cases, it is a single shard.

@@

To address the gap between pure \DDT and IOD's transactions, example data
storage and metadata services were created. These examples were not built for
performance but to determine the minimum functionality requirement for a
functional service. While these services were built without knowledge of how
IOD and DAOS operate, we came to similar functionality. The major difference is
that there are no dependencies between transactions that prevent visibility
should an older version be incomplete. This additional, intentional requirement
by IOD offers different functionality than \DDTns's example services. In the
case of \DDTns, the functionality is more minimal, but also avoids some of the
concerns outlined below.

@@

At a a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree. For IOD, it is a server-side tree. In both
cases, there is a master in charge of managing the transaction and a collection
of workers that aggregate into the master through second-level leaders. Beyond
that, there are some significant differences. Some of the different choices
made by IOD raise some possible concern.

@@

Fourth, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an evnironment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection.

@@

\subsection{Comparision to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guaranteed for when consensus will be
achieved without using slower synchronous calls. For the tight timing we wish
to support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments. The Intel/Lustre FastForward epoch
approach~\cite{barton:2013:fastforward,lombardi:2013:epochs} is discussed in
Section~\ref{lab:evaluation}.

Third, the server must offer a way to mark an operation as ``committed'' during
the ``commit'' phase of the two-phase commit.  Again, this may be implemented
in a transactional wrapper.  As an alternative, a scalable locking service like
Chubby~\cite{burrows:2006:chubby} might be employed.

Fourth, a second layer of coordination on the client side is introduced that
greatly increases the scalability by consolidating messages from clients into
unique sets prior to sending to the overall coordinator. A gossip
protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
purpose, but the delay of eventual consistency is strictly avoided with this
protocol to ensure guarantees at particular states in the code. For example, if
a failure occurs, the global understanding of the role of all processes is
required in order for effective communication to occur for operations like
creating sub-transactions or voting. In this case, the protocol can offer
stronger statements about consistency than these protocols offer.  These
features offer a way to easily scale the transaction protocol given the
guarantees we wish to offer.

@@

@@

We have demonstrated that synchronous two-phase commit transactions can have
low overhead and be used with roughly synchronous operations with little
additional overhead. While other use cases, particularly those that require a
greater degree of asynchrony, require a different approach, \DDT offers a
working solution for a wide variety of scenarios today.  The bigger question of
what sort of synchronization mechanism is appropriate for various use cases is
currently under investigation.

The evaluation of the fault detection and recovery are beyond scope for this
paper and are in preparation for presentation elsewhere.

>>>>

\begin{abstract}
The Fast Forward effort to define a next generation filesystem has made
admirable contributions in architecture and design. Formalizing the general
idea of data staging as burst buffers for the file system will help manage
the performance variability and offer the data processing opportunities
outside the main compute and file system. Adding a tranasctional mechanism to
manage faults and data visilibilty helps enable effective analytics without
having to work around the IO stack semantics.

While these and other their contributions are valuable, similar efforts made
elsewhere may offer attractive alternatives or differing semantics that would
yield a more feature rich environment with little to no additional overhead.
For example, the Doubly Distributed Transactions (\DDTns) protocol offers an
alternative approach for incorporating transactional semantics into the data
path. The PreDatA project examined how to get the best throughput for data
operators can offer some insights for refining the Burst Buffer concept.

This paper examines some of the choices made by the FastForward team and
compares them with other options and offers observations and suggestions based
on these other efforts.  This will include some non-core contributions, such
as some of the demonstration components generated while implementing \DDT,
to make suggestios that may help the next generation design for how the IO
stack works as a whole.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve the needs of an extreme scale platform. To
address these limitations, the US Department of Energy commissioned an effort
to develop a design and prototype for an IO stack suitable for the extreme
scale environment. This is a joint effort led by the Intel Lustre team, Los
Alamos National Laboratory, EMC, and the HDF Group. This team has developed a
specification set~\cite{fastforward:2014:docs} for a future IO stack to address
the identified challenges.

The basic architecture incorporates five layers. The top layer is a high level
IO library, such as the demonstration HDF-5 library. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers. Below the user API is an IO forwarding layer
that redirects IO calls from the compute nodes to the IO dispatching layer.
This IO forwarding layer is analogous to the function of the IO nodes in a
BlueGene machine. The next two layers have considerable functionality. The IO
dispatcher (IOD) serves as the primary storage interface for the IO stack and
is the only way to reach the persistent storage array in lower layers.

The core idea for IOD is to provide a way to manage the IO load that is
separate from the compute nodes and the storage array. Communication intensive
acitivies, such as two-phase, data sieving IO's data rearrangement process, can
be moved to the IOD layer offloading the communication load from the compute
nodes. IOD has three main purposes. First, if the optional burst buffer is
availble, it works as a fast cache absorbing write write operations for the
slower trickle out to the central storage array. It can also be used to
retrieve objects from the central storage array for more efficient read
operations and offer data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults. Third, data processing operations can be placed in the
IOD. This operations are inteded to offer data rearrangement, filtering, and
similar operations prior to it reaching the central storage array.

This offloads the requirement for collective two-phase data sieving
on the compute nodes to reorganize data. This technique has been proven
effective at reducing the total time for writing data due to fewer participants
involved in the communication patterns~\cite{lofstead:2011:nessie-staging}.
Beyond these broad items, there are many important details some of which are
examined in more detail below.

@@

The focus of this paper is primarily the IOD layer given the critical role it
has in the performance and functionality of the entire stack. Some of the key
features explored in this paper all have a focus on the IOD layer motivating
the focus of this examination.

@@

A second feature is the incorporation of a transaction like mechanism into the
IOD layer and a related epochs feature into the DAOS layer. This feature offers
a mechanism to avoid making incomplete data visible to other users and offer a
level of protection if the data can be stored on the optional SSDs. The
implications of how these concepts are designed are important to consider for
the final design. Because they control data visibility and manage faults for
output operations, they are key features.

The rest of the paper is organized as follows. Section~\ref{sec:burst}
discusses some of the features of incoporating burst buffers as designed and
suggests some considerations and alternatives for the next generation of this
project. Section~\ref{sec:transactions} discusses the transactions approach
offered in the IOD layer and the corresponding epochs in the DAOS layer. It
also offers a comparison to the \DDT system given the very similar high-level
design and motivating use case.  Section~\ref{sec:optimizations} discusses some
of the optimizations offered by the current design and how and when these
optimizations will work best. It also explores some of the dark corners that
may yield performance issues with the current design. Finally,
Section~\ref{sec:summary} disucsses the system overall with recommendations on
what design elements should be considered based on broader issues with current
HPC data centers.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any hardware support for this functionality. The
desired outcome of these initial studies is to motivate how such functionality
might be incorporated and the potential benefits.  Later, these concepts were
proposed to be incorporated as part of the IO
stack~\cite{bent:2012:challenges,bent:2012:burst-buffer}.  The current
FastForward IOD design recommends incorporating SSDs, but specifically lists
these devices as optional. Unfortunately, both incorporating burst buffers and
the use of SSDs in the IOD layer may be problematic.  First, with these burst
buffers being optional, the semantics of how IOD works must change to use DAOS
to store the data directly rather than storing them locally until explicitly
persisted by the user. The design of DAOS does not incorporate a high
transaction rate mechanism. Conversely, it assumes that it will only be
involved when persisting a completed transaction and only for a fraction of the
total transactions created. Further, with function shipping offering the
ability to change how the data is stored and arranged prior to it being written
to DAOS, this functionality may also be optional depending on the existence of
the burst buffer. Consider the important functionality of data rearrangement
and doing things like changing the fast array dimension.

One of the bigger concerns is the observation that the original data staging
proposals all used compute nodes while the newer proposals seek not only to
make them a fixed portion of the IO stack, but also shared across all machine
users. The PreDatA paper in particular examines the potential costs and
advantages of where to place operators similar to the IOD proposed function
shipping. There are two key takeaways from that work. First, placement matters.
Depending on the communication intensity vs. computation intensity, where
along the IO path to place the operation can matter significantly. Second, and
more importantly, the amount of time spent processing for the operators was
stretched to the point where it consumed nearly all of the time between IO
operations. The given ratios of compute processes to staging process examined
is representative for future extreme scale platforms. If anything, the ratios
offer more staging processes than IOD processes would be available.

By concentrating this functionality into the storage stack, two problems arise.
First, the amount of network bandwidth, IO bandwidth, and compute power
consumed for example operations from a single application is likely to
completely monopolize the IOD processes. Second, if space and time partitioning
is used instead, the functionality risks being too small to be useful. Third,
the hardware performance or SSDs is questionable. Current NAND-based flash
devices top out at around 400 MB/sec. The key spec that is missing from this
number is that 400 MB/sec is a measure of the fixed number of available IOPS
multipled by the block size. This represents the ideal streaming performance
possible. The problem is that it costs an IOP to read 1 byte or 1 block. It
costs 1 IOP to write a full block--usually. In some cases, it will cost 2 IOPs.
This accounts for the required pre-erase write prior to writing the block. In
the worst case, it can be 3 IOPS per write. This would be for a partial block
write (read the old block, erase, write the new block). One-third of 400
MB/sec, about 133 MB/sec, is well below the streaming performance of HDDs.
Granted, there is still rotational and seek latency to deal with for HDDs, but
the advantage for SSDs has evaporated and potentially turned into a penalty.

Given these features, the optionality and even incorporation of burst buffers
in the current design should be reconsidered. Much of the advanced, key
functionality proposed as they are currently designed ultimatey relies on the
existence of burst buffers to work. Further thought about how to have an IOD
layer both with and without a burst buffer is required before they can be
considered optional. As the design stands today, they are a required part of
the IOD layer for proper functioning. Unfortunately, it is not clear that they
can address the performance concerns they are intended to cover.

@@

The transaction mechanism manifests in two forms. At the IOD layer, they are
called transactions and are used to judge whether or not an output is complete
or not and control access through treating the transaction ID as a version
identifier. At the DAOS layer, they are called epochs and represent persisted
(durable) transactions from the IOD layer. Each of these offers different
functionality, but are connected as is explained below. How these differ from
the \DDT approach is also explored. While they are seemingly very different,
they use a very similar approach to solve the same problem.

\subsection{IOD Transactions}
To understand how transactions are used in the IOD layer, some terminology and
concepts must be explained first. At the coarsest grain level is a container.
Each container serves to host a transaction and contains a collection of
objects. Conceptually, containers corresponds to a file in a traditional file
system. The objects in each container represent different data within a file.
The three initially defined object types are key-value store, array shard,
and blob. The easiest way to understand these types is to evaluate these from
the perspective of an HDF-5 file, the initial user interface layer. The
key-value store represents a collection of attributes. The array shard
represents part of a potentially multi-dimensional array. It is refered to as a
shard because it is likely a small piece of a globally defined array. The blob
represents a byte stream of arbitrary contents.  The fundamental difference
between an array shard and a blob is that the array shard has metadata
identifying its portion within the global, logical space while the blob is
simply a 1-dimensional array of bytes that is not shared across IOD nodes.
Should an operation be deployed into IOD to manipulate an array within a
container, it would operate on the array shards rather than on blobs. Given
this context, the transactions come in two forms.

@@

\subsection{DAOS Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted copies
of a container. To simplify the mapping between an IOD transaction and the DAOS
epochs, the IOD transaction ID is the same as the epoch ID. The key difference
is that at the DAOS layer, some transaction (epoch) IDs will not be
represented.

\subsection{Metadata Management}
While a central design goal of the Fast Forward project to eliminate the
requireent for a metadata server as a core component, it was not fully
accomplished.  This idea is not new and has been explored as part of the Light
Weight File Systems project~\cite{oldfield:lwfs}. In LWFS, the metadata service
is explicitly limited to a user task with the storage layer limited to data
storage/retrieval, authorization, and authentication. This approach proved
workable. For the Fast Forward project, while a similar goal is desired, it is
not completely accomplished.

The purpose behind the goal is to eliminate the serialized bottleneck caused by
having a centralized metadata service. Much of the design of IOD with objects
and continers does work to eliminate these bottlenecks.  However, IOD does not
go fully to a no metadata service model. Instead, the IOD layer describes a
key-value metadata service it provides. This service maintains four different
items:

@@

While a pure no metadata model would be ideal for performance, this current
proposal has a few challenges. By providing this basic functionality, metadata
management is split between the user layer and the IOD layer. Consdier the
following items. First, there are no list of containers. This is closer to the
no metadata ideal, but given the rest of the functionality, it exports to the
user level a requirement to track what containers are in the file system.
Second, given the flattening that occurs between IOD and DAOS, the mapping from
IOD objects to DAOS objects is unlikely to be 1-to-1. This is likely a list
instead. Further, given the ability to re-stripe in the IOD layer, the list is
not a list of objects completely included in the IOD object, but instead
objects from which part of the data came from some listed DAOS object.  Third,
sharding IOD objects makes sense based on the compute process connection a
shard has with the complete object. Striping of an IOD object makes sense in
connection with the DAOS object(s) underneath, but striping within the IOD
layer is not discussed. Further, using striping from clients into the IOD layer
introduces a level of coordination that IOD seeks to avoid.  Fourth, given the
previous items, there is ambiguity with the maximum valid offset of the object.
This could be for the shard or for the collection of shard objects that
represent a distributed variable. Ideally, either a robust metadata management
system should be incorporated into the system, preferably separate from the IOD
layer. Alternatively, all of this functionality should be left to the user
layer. The only potential challenge is determining the container for a file
open command and what the transaction ID should be. This limited operation
should be a straightforward search using a hash or similar fast mechanism for
examining a list of container names and/or object IDs.

The \DDT project created a simplistic model for a metadata service that has
different features. It is described below and has been published
previously~\cite{lofstead:2012:txn-metadata}. Some of the major differences
include the following. First, \DDTns's metadata service has a way to get a list
of the objects in the metadata service. Since this was intended as a test
system to examine \DDT performance, we did not put the effort into a making it
a robust system. While it does not address having a file equivalence like the
container concept, it does introduce a way to figure out what is available.
Stored along with this object description is a list of the global dimensions
for array objects. One observtion made as part of creating this service is that
this is insufficient for many engineering codes that use non-regular meshes.
Second, for each object, a list of the equivalent to shards are kept with a
link back to the master object and the offset and size for each dimension for
that shard. This approach is not scalable because of the explosion in the
process count. Using a striping approach like IOD does is superior. Third,
there is no mapping to another layer with different object layout and counts.

Based on the lessons from the \DDT metadata service construction and the prior
experiments with LWFS, having a completely separate metadata service is
workable. Rather than making it a bottleneck in the IO path, it is another
service that users must interact with if they need those services. Otherwise,
it is 100\% out of the way. Users can manage everything by maintaining the
metadata including the list of objects themselves. However, there are drawbacks
to this 100\% client-side approach.

For the IOD proposal, the service needs some adjustments to be generally
useful.  First, there needs to be a way to query the list of available
containers. This will support an analysis application using data generated by
the simulation code. This use case is a stated motivating example for the Fast
Forward project. Second, the maximum transaction ID for a container is
mentioned as being maintained by the IOD metadata service, but when and how
this is done is never mentioned.  Third, the definition of an object needs to
be made more clear. In some cases, it seems that an object is a globally
distributed object is represented by a collection of shards. In other cases, it
is a single shard.

@@

The example metadata and data storage services created as part of the \DDT
project lack one of the key transaction management features used by IOD.
The major difference is that there are no dependencies between transactions
that prevent visibility should an older version be incomplete. This additional,
intentional requirement by IOD offers different functionality than \DDTns's
example services. In the case of \DDTns, the functionality is more minimal, but
also avoids some of the concerns outlined below.

@@

At a a high level, both \DDT and the IOD transactions have the same high-level
design. In both cases, a hierarchical model is employed. In the case of \DDTns,
it is a purely client-side tree using semi-synchronous messaging. The messaging
itself, in the current implementation, uses asynchronous MPI messages. The
synchronous component comes from the timeout mechanism used to detect faults.
It forces a level of coordiation and synchronization for the protocol. For IOD,
it is a server-side tree and fully asynchronous. In both cases, there is a
master in charge of managing the transaction and a collection of workers that
aggregate into the master through second-level leaders. Beyond that, there are
some significant differences. Some of the different choices made by IOD raise
some possible concern.

@@

Fourth, in the multi-leader model, using a count of client connections to
determine if transaction is complete is problematic.  First, should this be in
an evnironment where lost or repeated network messages are possible, then the
count could be wrong inappropriately. This is a general problem and unlikely
in the current environment, but is a concern on less integrated platforms.
The idea of a count precludes any tracking of expected messages from sources
introducing a potential hole. Second, the aggregation of completion messages
are based on the local aggregation point reporting to the transaction
leader the state.  There is no ability to detect or recover from a failure of
this IOD process that would report to the transaction leader. Third, because a
simple count of participating processes is sent with the beginning of the
transaction, each process is strictly limited to a single output operation.
This precludes a process writing multiple array shards, for example. To put
this in concrete terms, it means that a file could only contain a single,
globally distributed variable (at the root level of HDF5) OR an attribute per
process. If this count actually represents a begin/end transaction pair
instead, it suffers from a lack of accountability for missing object (shard)
writes. All it would indicate is that a process started and ended a transaction
connection. There is no validation that a process has done everything required
of it.

@@

\subsection{Comparision to Other Protocols}
Alternatives, such as Paxos~\cite{Lamport:1998:paxos} algorithms like
ZooKeeper~\cite{Hunt:2010:zookeeper}, suffer from two limitations making them
unsuitable for this environment. First, the distributed servers in Paxos
systems are all distributed copies of each other that eventually become
consistent. Given the scale we wish to address, a single node's memory is
unlikely to be able to hold all of the data necessary for many operations at
scale. They also do not have a guarantee for when consensus will be achieved
without using slower synchronous calls. For the tight timing we wish to
support, we need guarantees of when a consistent state has been achieved.
Second, these systems also all assume that updates are initiated from a single
client process rather than a parallel set of processes as is the standard in
HPC environments.

As an alternative, a scalable locking service like Chubby~\cite{burrows:2006:chubby} might be employed.

\DDT uses a second layer of coordination on the client side that greatly
increases the scalability by consolidating messages from clients into unique
sets prior to sending to the overall coordinator. A gossip
protocol~\cite{ganesh:2003:gossip-protocols} may appear sufficient for this
purpose, but the delay of eventual consistency is strictly avoided with this
protocol to ensure guarantees at particular states in the code. For example, if
a failure occurs, the global understanding of the role of all processes is
required in order for effective communication to occur for operations like
creating sub-transactions or voting. In this case, the protocol can offer
stronger statements about consistency than these protocols offer.  These
features offer a way to easily scale the transaction protocol given the
guarantees we wish to offer. The IOD approach does nothing to address these
concerns.

@@

THIS MUST BE RECONCILED WITH THE EARLIER TEXT THAT SAYS LARGELY THE SAME THING.

@@

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the concensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are glaring
oversights in the current documents. It is our undertanding that these will be
addressed in the next phase and we hope to help inform that effort with our
experiences.

We hope that the efforts made in the \DDT project and other efforts to explore
the requirements for this space, along with the analysis presented in this
paper will prove useful for the next phase of the Fast Forward project.
